---
title: "Maps"
author: "Diego"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Paquetes and files

```{r}
pacman::p_load(tidyverse, googlesheets4,ggridges,viridis,hrbrthemes, lubridate, gdata,dplyr,zoo,rgeoboundaries, raster,ggspatial,rnaturalearth,rgeos,leaflet,chirps, sf,zoo,rio,rnaturalearthdata,geodata,rgdal,terra,raster,tiff)
remotes::install_gitlab("dickoa/rgeoboundaries")  
```


# Mapas climaticos
## Ubicacion
```{r}
africa <- ne_countries(continent = "Africa", returnclass = "sf", scale = "medium")
namibia <- africa %>% filter(sovereignt=="Namibia")
africa %>%  
  ggplot()+
  geom_sf() + 
  geom_sf(data=namibia, fill="gray50")
```
## Maps of administrative 1 and 2
```{r}
namibia_adm1 <- geoboundaries(c("namibia"), "adm1")
zambezi <- namibia_adm1%>% filter(shapeName=="Zambezi Region")
namibia_adm2 <- geoboundaries(c("namibia"), "adm2")
zambezi_div<-namibia_adm2%>% filter(shapeName=="Kabe"|shapeName=="Kongola"|shapeName=="Linyandi"|shapeName=="Sibinda"|shapeName=="Katima Muliro Rural"|shapeName=="Katima Muliro Urban")


namibia  %>% 
ggplot() +
  geom_sf()+
  geom_sf(data= zambezi, fill="gray50")


namibia %>% 
ggplot() +
  geom_sf()+
  geom_sf(data= zambezi, fill="gray50")

```
```{r}
library(htmltools)

zambezi_div %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons(label = zambezi, weight=1, popup = ~htmlEscape(shapeID))
```
## Mapa con variables climaticas
```{r}
tmin_data <- getData(name = "worldclim", var = "tmin", res = 0.5,lon=24, lat=-18)
gain(tmin_data) <- 0.1
tmin_mean <- mean(tmin_data$tmin7_36)

tmin_mean_zam <- raster::mask(tmin_mean, as_Spatial(zambezi))
tmin_mean_zam_df <- as.data.frame(tmin_mean_zam, xy = TRUE, na.rm = TRUE)

```


```{r}
t_min_map<- tmin_mean_zam_df %>%
  ggplot(aes(x = x, y = y)) +
  geom_raster(aes(fill = layer)) +
  geom_sf(data = zambezi, inherit.aes = FALSE, fill = NA) +
  labs(
    title = "Mean monthly min temperatures in July in Zambezi",
    subtitle = "For the years 1970-2000"
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  scale_fill_gradient(
    name = "Temperature (°C)",
    low = "#FEED99",
    high = "#AF3301"
  )
  ggsave(last_plot(), file="Plots/tmin_map.tiff", 
       w=150, h=120, units="mm", dpi=300)

```
##Tmin in JUly in serondela
```{r}

tmin_mean_zam_df %>%
  ggplot(aes(x = x, y = y)) +
  geom_raster(aes(fill = layer))+
  xlab("Longitude")+
  ylab("Latitude")+
  xlim(c(24.9, 25.1))+
  ylim(c(-17.9, -17.7))+
  labs(
    title = "Tmin July Kabulabula"
  )
 scale_fill_gradient(
    name = "Tmin",
    low = "#FEED99",
    high = "#AF3301"
  )
```


## subir mapas de otras variables ambientales___altitude
```{r}
alt_data <- getData(name = "worldclim", var = "alt", res = 0.5,lon=24, lat=-18)
plot(alt_data)
alt_zam <- raster::mask(alt_data, as_Spatial(zambezi))
alt_zam_df <- as.data.frame(alt_zam, xy = TRUE, na.rm = TRUE)
alt_zam_df %>%
  ggplot(aes(x = x, y = y)) +
  geom_raster(aes(fill = alt_36)) +
  geom_sf(data = zambezi, inherit.aes = FALSE, fill = NA) +
  labs(
    title = "Altitude"
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  scale_fill_gradient(
    name = "ALtitude (mals)",
    low = "#FEED99",
    high = "#AF3301"
  )
``


##Altitude Kabulabula
`
```{r}
alt_zam_df %>%
  ggplot(aes(x = x, y = y)) +
  geom_raster(aes(fill = alt_36))+
  xlab("Longitude")+
  ylab("Latitude")+
  xlim(c(24.9, 25.1))+
  ylim(c(-17.9, -17.7))+
  labs(
    title = "Altitude Kabulabula"
  )
 scale_fill_gradient(
    name = "ALtitude in KAbulabula (mals)",
    low = "#FEED99",
    high = "#AF3301"
  )
```
## soil from ISDA https://gitlab.com/openlandmap/africa-soil-and-agronomy-data-cube
# K in ppm 0-20cm
```{r}
extent(zambezi)
crs(zambezi)



tif.cog = paste0("/vsicurl/https://s3.eu-central-1.wasabisys.com/africa-soil/layers30m/", 
              c("sol_log.k_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_bdr_m_30m_0..200cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_clay_tot_psa_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.c_tot_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.ca_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.mg_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.n_tot_ncs_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.oc_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.p_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.s_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_log.zn_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_ph_h2o_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_sand_tot_psa_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif",
              "sol_texture.class_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif"
              ))

sol30m = lapply(tif.cog, function(i){crop(rast(i), zambezi)})  
sol30m.sp = do.call(cbind, lapply(sol30m, function(i){as.data.frame(i)}))

str(sol30m.sp)

sol30m.sp$k_mehlich3_l1 = expm1(sol30m.sp$sol_log.k_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1 / 10)

pol.zam = rasterize(zambezi, sol30m[[1]])
sol30m.m = as(as(raster(pol.zam), "SpatialGridDataFrame"), "SpatialPixelsDataFrame")
sol30m.m$k_mehlich3_l1 = sol30m.sp$k_mehlich3_l1[sol30m.m@grid.index]

spplot(sol30m.m["k_mehlich3_l1"])


```

# p
```{r}

pol.v <- vect(zambezi)

tif.cog = paste0("/vsicurl/https://s3.eu-central-1.wasabisys.com/africa-soil/layers30m/",
              "sol_log.p_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1.tif"
              )

sol30m = lapply(tif.cog, function(i){crop(rast(i), pol.v)})  
sol30m.sp = do.call(cbind, lapply(sol30m, function(i){as.data.frame(i)}))

#Next, we need to back-transform values from log-scale to ppms:

sol30m.sp$p_mehlich3_l1 = expm1(sol30m.sp$sol_log.p_mehlich3_m_30m_0..20cm_2001..2017_africa_epsg4326_v0.1 / 10)


pol.r = rasterize(pol.v, sol30m[[1]])
sol30m.m = as(as(raster(pol.r), "SpatialGridDataFrame"), "SpatialPixelsDataFrame")
sol30m.m$p_mehlich3_l1 = sol30m.sp$p_mehlich3_l1[sol30m.m@grid.index]

spplot(sol30m.m["p_mehlich3_l1"])
```


## Upload maps from qgis project

```{r}

GDALinfo(mg_raster)
extent(mg_raster)
crs(mg_raster)
plot(mg_raster)

mg_raster <- raster(x = "Data/Mg_20_PPM.tif")
 t = readTIFF("Data/Mg_20_PPM.tif", as.is=TRUE)
```

## GRownd water drought risk
```{r}
gw_drought_risk<- readTIFF("Data/Groundwater_drought_risk_updated_final/GRoundwater_drought_risk_updated_final.tif")
plot(gw_drought_risk)
```








